Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2024-03-26 paper1_jiturra.Rmd
bonitos plots

```{r fig1, fig.dim=c(12,5),fig.cap='Bivariate relationships between Income Inequality, Social Class and Network homogeneity',warning=FALSE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

@@

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         # gini_disp=scale(gini_disp),
         )

homo_ineq <- dfreg %>%
  # filter(country3 != "ZAF") %>%
  group_by(country2,country3,dclass3res_V) %>%
  summarise_at(c("homclass3_V_res","egal","gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame()

homo_ineq$country <- countrycode::countrycode(sourcevar = homo_ineq$country2,origin = "iso3c",destination = "iso2c")

homo_ineq$dclass3res_V <- forcats::fct_rev(homo_ineq$dclass3res_V)
homo_ineq %>%
ggplot(aes(y=homclass3_V_res, x=gini_disp, group=dclass3res_V,fill=dclass3res_V)) +
  geom_text(label=as.character(homo_ineq$country),size=3) +
  geom_smooth(method = "lm",fullrange=TRUE, color="black",linetype = 2,size=1,se = FALSE) +
  facet_wrap(~dclass3res_V) +
  xlab("Income Inequality")+
  ylab("Network homogeneity") +
  # labs(title = "Class homogeneity and Gini (post-taxes)") +
  scale_fill_manual(values=c("#CCCCCC","#989898","#323232"))+
  scale_color_manual(values=c("#CCCCCC","#989898","#323232")) +
  theme(legend.position = "none",legend.title = element_blank())

@@

>>>>

```{r fig1, fig.dim=c(10,5),fig.cap='Cross-country comparison of Redistributive preferences and Social Class ',warning=FALSE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,ggplot2,haven,ggrepel,gridExtra)  
rm(list=ls())

@@

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  gdppercapita,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         # gini_disp=scale(gini_disp),
         )

df_mean <- 
dfreg %>% 
  group_by(country3) %>% 
  summarise(egal2_c=mean(egal),homclass_c=mean(homclass3_V_res),gini_disp=mean(gini_disp))

df_region <- 
dfreg %>% 
  group_by(region,country3) %>% 
  summarise(n())

df_full <- 
dfreg %>% 
  group_by(dclass3res_V) %>% 
  summarise(egal2=mean(egal),homclass=mean(homclass3_V_res),homclass_c=mean(dfreg$homclass3_V_res),egal2_c=mean(dfreg$egal), country3="Sample")

df_plot <- 
dfreg %>% 
  group_by(country3,dclass3res_V) %>% 
  summarise(egal2=mean(egal),homclass=mean(homclass3_V_res)) %>% 
  left_join(df_mean) %>%
  left_join(df_region) 
# df_plot <- 
# df_plot %>% 
#   bind_rows(df_full)

df_plot$dclass3res_V <- factor(df_plot$dclass3res_V,levels = levels(df_plot$dclass3res_V),labels = c("Service Class","Intermediate Class","Working Class"))

df_plot %>% 
ggplot()+
  geom_hline(yintercept = mean(dfreg$egal),linetype=2)+
  geom_point(aes(x = reorder(country3, egal2_c),y = egal2,shape=dclass3res_V,color=dclass3res_V),size=3, alpha=2) +
  geom_point(aes(x = reorder(country3, egal2_c),y = egal2_c),alpha=1,shape=4,size=2) +
  scale_shape_manual(values = c(17, 15, 16)) +
  scale_color_manual(values = c("#000000","#666666","#999999")) +
  labs(x=NULL,y="Redistributive Preferences") +
  theme(legend.position = c(0.90, 0.15),
        axis.text.x=element_text(angle=45,hjust = 1,size=8),
        legend.title = element_blank())

# df_plot %>%
# ggplot()+
#   geom_point(aes(x = reorder(country3, homclass_c),y = homclass,shape=dclass3res_V,color=dclass3res_V),size=3, alpha=2) +
#   geom_hline(yintercept = mean(dfreg$homclass3_V_res),linetype=2)+
#   geom_point(aes(x = reorder(country3, homclass_c),y = homclass_c),alpha=1,shape=4,size=2) +
#   scale_shape_manual(values = c(17, 15, 16)) +
#   scale_color_manual(values = c("#000000","#666666","#999999")) +
#   labs(x=NULL,y="Network homogeneity") +
#   theme(legend.position = c(0.90, 0.1),
#         axis.text.x=element_text(angle=45,hjust = 1,size=10),
#         legend.title = element_blank())
```

```{r fig2, fig.dim=c(10,5),fig.cap='Bivariate relationships between Income Inequality, Network homogeneity and Redistributive preferences',warning=FALSE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,ggplot2,haven,ggrepel,gridExtra)  
rm(list=ls())

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  gdppercapita,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         # gini_disp=scale(gini_disp),
         )
homo_ineq <- dfreg %>%
  # filter(country3 != "HUN") %>%
  group_by(country2,country3) %>%
  summarise_at(c("homclass3_V_res","egal","gini_disp","d10d1"), mean, na.rm = TRUE) %>% 
  as.data.frame()

# Plot Homogeneity and redistributive preferences
homo_ineq$country <- countrycode::countrycode(sourcevar = homo_ineq$country2,origin = "iso3c",destination = "iso2c")

fig_egal_homo <- 
homo_ineq %>%  
ggplot(aes(y=egal, x=homclass3_V_res)) +
  geom_smooth(method = "lm",fullrange=TRUE, color="black",linetype = 2,size=1,se = T) +
  geom_text_repel(label=as.character(homo_ineq$country)) +
  xlab("Network homogeneity")+
  ylab("Redistributive preferences") +
  labs(caption =paste("r = ",round(cor(homo_ineq$homclass3_V_res,
                                     homo_ineq$egal, use='complete.obs'),digits = 2)),tag = "A")
fig_homo_gini<- 
homo_ineq %>%  
ggplot(aes(y=homclass3_V_res, x=gini_disp)) +
  geom_smooth(method = "lm",fullrange=TRUE, color="black",linetype = 2,size=1,se = T) +
  # geom_text() +
  geom_text_repel(label=as.character(homo_ineq$country)) +
  xlab("Income Inequality") +
  ylab("Network homogeneity") +
  labs(caption =paste("r = ",round(cor(homo_ineq$homclass3_V_res,
                                     homo_ineq$gini_disp, use='complete.obs'),digits = 2)),tag = "B")
grid.arrange(fig_egal_homo,fig_homo_gini, nrow = 1)
```

@@

```{r oldcode2, eval=FALSE, fig.cap='Bivariate relationships between Income Inequality, fig.dim=c(12,5), include=FALSE, Social Class and Network homogeneity',warning=FALSE}
homo_ineq <- dfreg %>%
  # filter(country3 != "ZAF") %>%
  group_by(country2,country3,dclass3res_V) %>%
  summarise_at(c("homclass3_V_res","egal","gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame()

homo_ineq$country <- countrycode::countrycode(sourcevar = homo_ineq$country2,origin = "iso3c",destination = "iso2c")
homo_ineq$dclass3res_V <- forcats::fct_rev(homo_ineq$dclass3res_V)
homo_ineq %>%
ggplot(aes(y=homclass3_V_res, x=gini_disp, group=dclass3res_V,fill=dclass3res_V)) +
  geom_text_repel(label=as.character(homo_ineq$country)) +
  geom_smooth(method = "lm",fullrange=TRUE, color="black",linetype = 2,size=1,se = FALSE) +
  facet_wrap(~dclass3res_V) +
  xlab("Income Inequality")+
  ylab("Network homogeneity") +
  # labs(title = "Class homogeneity and Gini (post-taxes)") +
  scale_fill_manual(values=c("#CCCCCC","#989898","#323232"))+
  scale_color_manual(values=c("#CCCCCC","#989898","#323232")) +
  theme(legend.position = "none",legend.title = element_blank())
